Eleven years of student replication projects provide evidence on the correlates of replicability in psychology

### 2023-03-30 paper/manuscript.Rmd
work on discussion

---
title: "Correlates of replicability: evidence from student replication projects"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: Maya Mathur
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251MA.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: dmk-format.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt

@@

# TODO list
- BETTER title
- BETTER abstract
- need to do supplement with the stuff we preregistered (sensitivity analysis, all models) -- waiting on finalized stats predictors
- many citations everywhere
- long list of things to gesture at in intro
- link to open stuffs
- ** verify code for ES ** with someone

# Citation slush pile

@@

@@

@@

## current work
Prior literature, such as @frank2012 and @hawkins have advocated using research methods classes to conduct replications, both for the education and scientific value.

@@

@@

Across the 176 studies, the overall subjective replication rate was `r round(pluck(sub_rep_rate,1))`%. `r round(pluck(pred_int_rate,1))`% of the studies had replication outcomes within the prediction interval of the original outcome. The median p_original value was `r round(pluck(porig_rate, 1),2)`. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

# discussion

blah blah we did a thing and it generally aligns with prior results. Overall replication rate is consistent with prior work, and we find cog greater than soc. However, by looking at a greater range of pr

[who knows where this paragraph should go] Replications are one way of approximating whether an effect in the target paper is true, and how likely results are to replicate in some sort of platonic ideal world. There is no one answer here; replication projects measure something else that could be treated as an approximation. However, what a replication is really measuring is how likely it to get that effect given some conditions. Which the right conditions are is a potential point of contention (see discussion around closeness and sample size, power etc).  Here, we are explicit in what sample of replicators and replication conditions were are sampling, and thus what we can generalize to. We are estimating how likely replication is when done online by a graduate student, under constraints that are typical for graduate students (limited time, limited budget). As much of scientific process is in fact performed by graduate students, we think this in itself is an interesting question. It contrasts with questions like how likely something is to replicate when performed by an expert with a large budget (perhaps this is the right framing on Camerer) or when performed with extensive feedback from the original authors, etc.

circle back to the idea that student replications (either as class or before extension into their own work) would be valuable

>>>>

---
title: "Eleven years of student replication projects provide evidence on the correlates of replicability in psychology"
author:
  - name: Veronica Boyce
    affiliation: Stanford
    footnote:
      - corresp
  - name: Maya Mathur
    affiliation: Stanford
  - name: Michael C. Frank
    affiliation: Stanford
address:
  - code: Stanford
    address: Stanford University
footnote:
  - code: corresp
    text: "Corresponding author. Email: vboyce@stanford.edu"
bibliography: ["251MA.bib"] # Replace with one or more of your own bibtex files. Better BibTeX for Zotero is your friend
csl: dmk-format.csl # Use any CSL style. See https://www.zotero.org/styles for a good list. Ignored if citation_package: natbib
link-citations: TRUE
output:
  bookdown::pdf_document2:
    toc: FALSE
    keep_tex: TRUE
    template: generic_article_template.tex
    #md_extensions: "-autolink_bare_uris"
    number_sections: TRUE
    citation_package: default # Can also be "natbib"
lang: en # Main document language in BCP47 format
geometry: "margin=25mm"
papersize: a4
#linestretch: 2 # for double spacing
endfloat: FALSE # Set to TRUE to turn on latex endfloat package to place figures and tables at end of document
# endfloatoption: # See endfloat documentation for more possibilities
#   - tablesfirst # Default
#   - nomarkers # Default
numberlines: FALSE
authblk: TRUE # FALSE = author affiliations in footnotes; TRUE = author affiliations in a block below author names
footnotehyper: FALSE # TRUE will give you enhanced table footnote capabilities. Set to FALSE to be able to use French blocks. Needed due to what appears to be a latex bug.
urlcolor: blue
linkcolor: blue
citecolor: blue
graphics: TRUE # Needed to be able to include images
tables: TRUE # Needed to be able to include tables
# fancyhdr:
#   first:
#     #headleft: "REPORT-NO-XXXX"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 0pt
#     #footleft: A left foot
#     footrulewidth: 0pt
#   subsequent:
#     #headleft: "NEXT-PAGE-HEADER-LEFT"
#     headright: "Kaplan et al. (2021)"
#     headrulewidth: 1pt
#     footrulewidth: 0pt

@@

# TODO list
- BETTER abstract
- many citations everywhere
- long list of things to gesture at in intro
- link to open stuffs
- ** verify code for ES ** with someone
- redo intro
- redo results
- redo discussion
- fill in todos

# Text slush pile 
[who knows where this paragraph should go] Replications are one way of approximating whether an effect in the target paper is true, and how likely results are to replicate in some sort of platonic ideal world. There is no one answer here; replication projects measure something else that could be treated as an approximation. However, what a replication is really measuring is how likely it to get that effect given some conditions. Which the right conditions are is a potential point of contention (see discussion around closeness and sample size, power etc).  Here, we are explicit in what sample of replicators and replication conditions were are sampling, and thus what we can generalize to. We are estimating how likely replication is when done online by a graduate student, under constraints that are typical for graduate students (limited time, limited budget). As much of scientific process is in fact performed by graduate students, we think this in itself is an interesting question.

circle back to the idea that student replications (either as class or before extension into their own work) would be valuable 
# Citation slush pile

@@

Anderson, C., Bahnik, S., Barnett-Cowan, M., Bosco, F., Chandler, J., Chartier, C., & otherss. (2016). Response to comment on “estimating the reproducibility of psychological science.” Science, 351(6277), 1037–1037.

Gilbert, D. T., King, G., Pettigrew, S., & Wilson, T. D. (2016). Comment on “estimating the reproducibility of psychological science.” Science, 351(6277), 1037–1037.

Schmidt, S. (2009). Shall we really do it again? The powerful concept of replication is neglected in the social sciences. Review of General Psychology, 13, 90–100.

Simonsohn, U. (2015). Small telescopes: Detectability and the evaluation of replication results. Psychol. Sci., 26(5), 559–569.

Gelman, A. (2018). Don’t characterize replications as successes or failures. Behavioral and Brain Sciences, 41.

Mathur, M. B., & VanderWeele, T. J. (2020a). New statistical metrics for multisite replication projects. J. R. Stat. Soc. Ser. A Stat. Soc., 183(3), 1145–1166.

pedagogy for replication in intro: frank & saxe, hawkins et al., quintana nature paper, scarcity PNAS pedagogy paper,

Collaborative replications and education project (CREP)
JE Grahe, MJ Brandt, H IJzerman, N Legate, J Wagge, YJ Weisberg, BJ Wiggins, Retrieved from osf. io/wfc6u, 2016 - Cited by 29

https://psycnet.apa.org/record/2022-15361-001
https://journals.sagepub.com/doi/pdf/10.1177/1475725719868149
https://content.iospress.com/articles/education-for-information/efi190308

https://psycnet.apa.org/fulltext/2018-11647-005.html

@@

replication is important (but not everything) - good starting place
 * prereq of theory building that the results of interest "can be observed consistently" 
 * but not everything (can still have other problems)

replication rates are low in psychology (see studies)

big open question: why

experimental and correlational approaches to answering it

experimental is hard - protzko, ML5 etc. but they are expensive and you can’t do them for many research questions

correlational has been pursued extensively, but is overfit to small data (prediction markets / ML models / regression of all types)

our contribution: provide more data!
using pedagogical replications, which are good because they’re p(build on)
slightly different estimand but maybe a better one?

@@

## current work

one strategy for the pedagogical bit at end of intro: stress “this is what people want to do, that is, pick up a project that they care about (not a random sample of the literature) and see if they can ‘get it to work’” - and “subjective replication success is our primary measure for the same reason….”

can even cite the “science is not a signal detection problem” paper on not randomly sampling from the literature.

Prior approaches to replicability have focused on a potentially problematic estimand: the probability of a finding in the literature being somehow truly replicable. Critics have pointed out that "true" replicability may not be possible to estimate outside of a specific sample (van bavel etc.) or even time period (ramscar).

Further, the methods for estimating this quantity have been theoretically problematic. Sampling schemes for prior work typically does not reflect a random sample from the literature -- and even if they did, arguably a literature will succeed if useful discoveries come out of it, not if random findings are true ("Science is not signal detection" ref).

In contrast, we have pursued a different estimand: the probability that a researcher, on selecting a finding of interest from the literature, can successfully achieve a result satisfactorily close enough to the original that they feel that they can build on it in their own work, with all the necessary compromises to the methods and sample of the original that may be required by the constraints of the situation. SPELL OUT SAMPLING DIFFERENCE AND SUCCESS DIFFERENCE.

We estimate this quantity in the context of a particular pedagogical situation: that of a graduate student entering a doctoral program in psychology...

Prior literature, such as @frank2012 and @hawkins have advocated using research methods classes to conduct replications, both for the education and scientific value.

@@

TODO discuss more about why this one and what other measure we don't choose are

@@

Across the 176 studies, the overall subjective replication rate was `r round(pluck(sub_rep_rate,1))`%. `r round(pluck(pred_int_rate,1))`% (N=FOOBAR) of the studies had replication outcomes within the prediction interval of the original outcome. The median p_original value was `r round(pluck(porig_rate, 1),2)`. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

# Discussion

The replication rate in psychology is to the best estimates, somewhere around 50%, which is quite low. Many reasons for this low replication rate have been hypothesized, but studying it either experimentally or correlationally requires doing a number of replication experiments, which can be a lot of work. Here we took advantage of 11 years of graduate student replication projects to look at correlational predictors of replication in a previously-unused dataset.

In line with previous replications, we found a FOOBAR replication rate, with some studies showing effect sizes similar to the original and others much smaller. Some individual correlates of replicability that stood out included within-subjects designs, work in the subfield of cognitive psychology, and the original and replication both using online samples. TODO other possible things

Many of these correlates interrelate with one another, and we are still limited in our sample, so models with multiple predictors have a lot of uncertainty around the effects of each predictor.

We are explicit in our goal of estimating how likely it is that a first-year graduate student, attempting to replicate a study of interest to them, gets replication results which are consistent enough with the original results that the student could build on the study in their own work. We can't estimate some platonic probability of truth for a study and neither can any replication study. Instead, we can estimate how likely studies are to get results of different levels of closeness in different circumstances. Much of the discussion around whether things should count as replication successes, or failures, or not replications at all is really about what the right thresholds are for closeness of results, and what the right circumstances are for a replication. Rather than try to find the one right answer to whether a direct replication was done in a certain right way, we think it's useful to explicitly label what the estimand really was: how likely it is to get the same results (under whatever metric) given circumstances like those used for the replication.

In our case, our estimand is about how well first-year graduate students will do at the replication, which takes into account the limited time, limited budget, and limited experience. We think this is an important estimand, as much of the work of psychology is done by trainees, in circumstances like these. Thus, if some studies have delicate results that require large samples and very exact methods to achieve, they may not replicate under normal resources.

Other replication projects have other estimands: @camerer2018 seems to ask something closer to how likely studies are to replicate when preformed by an expert with a large budget, and @ML5 TODOCITE asks how likely certain studies are to replicate when performed with extensive feedback from the original authors.

We do not interpret our results as saying that all non-replications were false positives (presumably some would be replicable under different implementations and budgets and others would not). There are many possible reasons for the non-replications in this sample. In some cases, it seemed that the problem may have been with the replication: for instance, if there were too few participants, or if there were high levels of wrong answers on attention checks, or participants speeding through without attention checks. For these cases, there was a clear next attempt that a student could make if they wanted to get the replication to work. In other cases, there might have been a priori reasons to distrust the original study results, such as exclusion criteria that seemed to be post-hoc, or a three-way interaction effect on a small sample (CITE THAT THIS IS SKETCH). In yet other cases, it's unclear why the results failed to replicate.

[somehow transition here] Pedagogy is important for open science. It's one thing to require or incentivize scientists to use open science practices and conduct replicable and reproducible research, but using the right tools and workflows to do open science is something that has to be learned. Teaching it in the classroom addresses the knowing how point at the beginning and shows students how to have open science practices integrated in to their science at the beginning, before other habits can ossify. Doing replications give students the motivation to care about open science, as they see how much easier it is to implement the study with open materials versus the study where they have to make guesses about the study instructions from the methods section. In presenting work with classmates, students see that there is variation in how well studies replicate, with some replicating very cleanly and others not at all. This sort of first hand experience teaches that not everything they read in the literature may just work if done again.

Our results are limited by the number and quality of the studies we included. These studies are not necessarily representative of the literature as a whole. TODO what are the limitations we want?

TODO quick wrap up

### 2023-03-30 paper/manuscript.tex
work on discussion

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  english,
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Correlates of replicability: evidence from student replication projects},
  pdflang={en},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=25mm]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\usepackage{footnote} % For some unknown reason, footnotehyper clashes with French
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% dont indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\title{Correlates of replicability: evidence from student replication projects}

@@

\begin{itemize}
\tightlist
\item
  BETTER title
\item
  BETTER abstract
\item
  need to do supplement with the stuff we preregistered (sensitivity analysis, all models) -- waiting on finalized stats predictors
\item
  many citations everywhere
\item
  long list of things to gesture at in intro
\item
  link to open stuffs
\item
  ** verify code for ES ** with someone
\end{itemize}

\hypertarget{citation-slush-pile}{%
\section{Citation slush pile}\label{citation-slush-pile}}

@@

@@

@@

@@

As a complement, we also used two statistical measure of replication on the subset of the data where they were computable (134 cases, see Figure \ref{fig:prisma}). We measured p-original, the p-value on the null hypothesis that the original and replication statistics are from the same distribution, as a continuous variable, and we also determined whether the replication statistic fell within the prediction interval of the original statistic (\protect\hyperlink{ref-errington2021}{Errington et al. 2021}).

@@

Across the 176 studies, the overall subjective replication rate was 49\%. 46\% of the studies had replication outcomes within the prediction interval of the original outcome. The median p\_original value was 0.03. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

\hypertarget{discussion}{%
\section{discussion}\label{discussion}}

blah blah we did a thing and it generally aligns with prior results. Overall replication rate is consistent with prior work, and we find cog greater than soc. However, by looking at a greater range of pr

{[}who knows where this paragraph should go{]} Replications are one way of approximating whether an effect in the target paper is true, and how likely results are to replicate in some sort of platonic ideal world. There is no one answer here; replication projects measure something else that could be treated as an approximation. However, what a replication is really measuring is how likely it to get that effect given some conditions. Which the right conditions are is a potential point of contention (see discussion around closeness and sample size, power etc). Here, we are explicit in what sample of replicators and replication conditions were are sampling, and thus what we can generalize to. We are estimating how likely replication is when done online by a graduate student, under constraints that are typical for graduate students (limited time, limited budget). As much of scientific process is in fact performed by graduate students, we think this in itself is an interesting question. It contrasts with questions like how likely something is to replicate when performed by an expert with a large budget (perhaps this is the right framing on Camerer) or when performed with extensive feedback from the original authors, etc.

circle back to the idea that student replications (either as class or before extension into their own work) would be valuable

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  english,
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Eleven years of student replication projects provide evidence on the correlates of replicability in psychology},
  pdflang={en},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=25mm]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\usepackage{footnote} % For some unknown reason, footnotehyper clashes with French
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% dont indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\title{Eleven years of student replication projects provide evidence on the correlates of replicability in psychology}

@@

\begin{itemize}
\tightlist
\item
  BETTER abstract
\item
  many citations everywhere
\item
  long list of things to gesture at in intro
\item
  link to open stuffs
\item
  ** verify code for ES ** with someone
\item
  redo intro
\item
  redo results
\item
  redo discussion
\item
  fill in todos
\end{itemize}

\hypertarget{text-slush-pile}{%
\section{Text slush pile}\label{text-slush-pile}}

{[}who knows where this paragraph should go{]} Replications are one way of approximating whether an effect in the target paper is true, and how likely results are to replicate in some sort of platonic ideal world. There is no one answer here; replication projects measure something else that could be treated as an approximation. However, what a replication is really measuring is how likely it to get that effect given some conditions. Which the right conditions are is a potential point of contention (see discussion around closeness and sample size, power etc). Here, we are explicit in what sample of replicators and replication conditions were are sampling, and thus what we can generalize to. We are estimating how likely replication is when done online by a graduate student, under constraints that are typical for graduate students (limited time, limited budget). As much of scientific process is in fact performed by graduate students, we think this in itself is an interesting question.

circle back to the idea that student replications (either as class or before extension into their own work) would be valuable
\# Citation slush pile

@@

Anderson, C., Bahnik, S., Barnett-Cowan, M., Bosco, F., Chandler, J., Chartier, C., \& otherss. (2016). Response to comment on ``estimating the reproducibility of psychological science.'' Science, 351(6277), 1037--1037.

Gilbert, D. T., King, G., Pettigrew, S., \& Wilson, T. D. (2016). Comment on ``estimating the reproducibility of psychological science.'' Science, 351(6277), 1037--1037.

Schmidt, S. (2009). Shall we really do it again? The powerful concept of replication is neglected in the social sciences. Review of General Psychology, 13, 90--100.

Simonsohn, U. (2015). Small telescopes: Detectability and the evaluation of replication results. Psychol. Sci., 26(5), 559--569.

Gelman, A. (2018). Don't characterize replications as successes or failures. Behavioral and Brain Sciences, 41.

Mathur, M. B., \& VanderWeele, T. J. (2020a). New statistical metrics for multisite replication projects. J. R. Stat. Soc. Ser. A Stat. Soc., 183(3), 1145--1166.

pedagogy for replication in intro: frank \& saxe, hawkins et al., quintana nature paper, scarcity PNAS pedagogy paper,

Collaborative replications and education project (CREP)
JE Grahe, MJ Brandt, H IJzerman, N Legate, J Wagge, YJ Weisberg, BJ Wiggins, Retrieved from osf. io/wfc6u, 2016 - Cited by 29

\url{https://psycnet.apa.org/record/2022-15361-001}
\url{https://journals.sagepub.com/doi/pdf/10.1177/1475725719868149}
\url{https://content.iospress.com/articles/education-for-information/efi190308}

\url{https://psycnet.apa.org/fulltext/2018-11647-005.html}

@@

replication is important (but not everything) - good starting place
* prereq of theory building that the results of interest ``can be observed consistently''
* but not everything (can still have other problems)

replication rates are low in psychology (see studies)

big open question: why

experimental and correlational approaches to answering it

experimental is hard - protzko, ML5 etc. but they are expensive and you can't do them for many research questions

correlational has been pursued extensively, but is overfit to small data (prediction markets / ML models / regression of all types)

our contribution: provide more data!
using pedagogical replications, which are good because they're p(build on)
slightly different estimand but maybe a better one?

@@

one strategy for the pedagogical bit at end of intro: stress ``this is what people want to do, that is, pick up a project that they care about (not a random sample of the literature) and see if they can `get it to work'\,'' - and ``subjective replication success is our primary measure for the same reason\ldots.''

can even cite the ``science is not a signal detection problem'' paper on not randomly sampling from the literature.

Prior approaches to replicability have focused on a potentially problematic estimand: the probability of a finding in the literature being somehow truly replicable. Critics have pointed out that ``true'' replicability may not be possible to estimate outside of a specific sample (van bavel etc.) or even time period (ramscar).

Further, the methods for estimating this quantity have been theoretically problematic. Sampling schemes for prior work typically does not reflect a random sample from the literature -- and even if they did, arguably a literature will succeed if useful discoveries come out of it, not if random findings are true (``Science is not signal detection'' ref).

In contrast, we have pursued a different estimand: the probability that a researcher, on selecting a finding of interest from the literature, can successfully achieve a result satisfactorily close enough to the original that they feel that they can build on it in their own work, with all the necessary compromises to the methods and sample of the original that may be required by the constraints of the situation. SPELL OUT SAMPLING DIFFERENCE AND SUCCESS DIFFERENCE.

We estimate this quantity in the context of a particular pedagogical situation: that of a graduate student entering a doctoral program in psychology\ldots{}

@@

As a complement, we also used two statistical measure of replication on the subset of the data where they were computable (138 cases, see Figure \ref{fig:prisma}). We measured p-original, the p-value on the null hypothesis that the original and replication statistics are from the same distribution, as a continuous variable, and we also determined whether the replication statistic fell within the prediction interval of the original statistic (\protect\hyperlink{ref-errington2021}{Errington et al. 2021}).

@@

Across the 176 studies, the overall subjective replication rate was 49\%. 45\% (N=FOOBAR) of the studies had replication outcomes within the prediction interval of the original outcome. The median p\_original value was 0.03. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The replication rate in psychology is to the best estimates, somewhere around 50\%, which is quite low. Many reasons for this low replication rate have been hypothesized, but studying it either experimentally or correlationally requires doing a number of replication experiments, which can be a lot of work. Here we took advantage of 11 years of graduate student replication projects to look at correlational predictors of replication in a previously-unused dataset.

In line with previous replications, we found a FOOBAR replication rate, with some studies showing effect sizes similar to the original and others much smaller. Some individual correlates of replicability that stood out included within-subjects designs, work in the subfield of cognitive psychology, and the original and replication both using online samples. TODO other possible things

Many of these correlates interrelate with one another, and we are still limited in our sample, so models with multiple predictors have a lot of uncertainty around the effects of each predictor.

We are explicit in our goal of estimating how likely it is that a first-year graduate student, attempting to replicate a study of interest to them, gets replication results which are consistent enough with the original results that the student could build on the study in their own work. We can't estimate some platonic probability of truth for a study and neither can any replication study. Instead, we can estimate how likely studies are to get results of different levels of closeness in different circumstances. Much of the discussion around whether things should count as replication successes, or failures, or not replications at all is really about what the right thresholds are for closeness of results, and what the right circumstances are for a replication. Rather than try to find the one right answer to whether a direct replication was done in a certain right way, we think it's useful to explicitly label what the estimand really was: how likely it is to get the same results (under whatever metric) given circumstances like those used for the replication.

In our case, our estimand is about how well first-year graduate students will do at the replication, which takes into account the limited time, limited budget, and limited experience. We think this is an important estimand, as much of the work of psychology is done by trainees, in circumstances like these. Thus, if some studies have delicate results that require large samples and very exact methods to achieve, they may not replicate under normal resources.

Other replication projects have other estimands: Camerer et al. (\protect\hyperlink{ref-camerer2018}{2018}) seems to ask something closer to how likely studies are to replicate when preformed by an expert with a large budget, and (\protect\hyperlink{ref-ML5}{\textbf{ML5?}}) TODOCITE asks how likely certain studies are to replicate when performed with extensive feedback from the original authors.

We do not interpret our results as saying that all non-replications were false positives (presumably some would be replicable under different implementations and budgets and others would not). There are many possible reasons for the non-replications in this sample. In some cases, it seemed that the problem may have been with the replication: for instance, if there were too few participants, or if there were high levels of wrong answers on attention checks, or participants speeding through without attention checks. For these cases, there was a clear next attempt that a student could make if they wanted to get the replication to work. In other cases, there might have been a priori reasons to distrust the original study results, such as exclusion criteria that seemed to be post-hoc, or a three-way interaction effect on a small sample (CITE THAT THIS IS SKETCH). In yet other cases, it's unclear why the results failed to replicate.

{[}somehow transition here{]} Pedagogy is important for open science. It's one thing to require or incentivize scientists to use open science practices and conduct replicable and reproducible research, but using the right tools and workflows to do open science is something that has to be learned. Teaching it in the classroom addresses the knowing how point at the beginning and shows students how to have open science practices integrated in to their science at the beginning, before other habits can ossify. Doing replications give students the motivation to care about open science, as they see how much easier it is to implement the study with open materials versus the study where they have to make guesses about the study instructions from the methods section. In presenting work with classmates, students see that there is variation in how well studies replicate, with some replicating very cleanly and others not at all. This sort of first hand experience teaches that not everything they read in the literature may just work if done again.

Our results are limited by the number and quality of the studies we included. These studies are not necessarily representative of the literature as a whole. TODO what are the limitations we want?

TODO quick wrap up
